---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pythonR
`pythonR` includes helper functions for using Python in R. It is essentially a wrapper of the `reticulate` R package, and can helpful in an R package development or project environment where you want more readable code for setting up a python environment and installing any necessary python dependencies.

It may decrease some of the flexibility offered via `reticulate` for setting up virtual or conda environments (mostly in ignoring the `RETICULATE_PYTHON` python path), but it allows you to reproducibly set up the same environment each session, and install all python packages in a single call.


You begin by setting up a python environment (conda or virtual) with the required imports you will need:
```{r}
 ## With a conda environment (the default) ##
 # much faster loading time after you've installed the packages once
 py_env <- setup_py_env(py_pkgs = c("pandas", "numpy", "scipy"))
 py_env

 # Put the conda environment under a specific name if you want to store multiple
 # note: you must restart your R session if you've already generated an environment
 py_env <- setup_py_env(
   py_pkgs = c("pandas", "numpy", "scipy"),
   conda_name = PYTHON_R_ENV,
   update = TRUE
 )

 ## Using a virtual environment ##
 py_env <- setup_py_env(py_pkgs = c("pandas", "numpy", "scipy"), virtual_env = TRUE)
 py_env

 # shutdown virtual environment
 shutdown_virtual_env(py_env$env_name)

```

Note that you can always install packages to the environment you created at a later time:
```{r}
 # Installing a package after setup (works with both environment types)
 setup_py_env(py_pkgs = c("pandas", "numpy"))
 install_py_pkgs(py_pkgs = c('scipy'))
```


Additional helper functions are provided for ensuring required python packages are installed or importing them at various stages within an R function/script. A simple example is provided below, but see `?pythonR_examples` for fully fleshed out examples and more details.


```{r}
my_py_function <- function(){
  # Check that python packages are installed
  required_py_pkgs <- c('pandas', 'scipy', 'numpy')
  checkmate::assert_true(check_py_pkgs_installed(required_py_pkgs))
  
    # import main python modules
  import_main_py()

  # import required python packages
  import_py_pkgs(py_pkgs = required_py_pkgs)
  
  # source python script or load python functions into R
  
  # do something...
}
```


```{r}
# Set up a python environment once per R session
py_env <- setup_py_env(py_pkgs = c("pandas", "numpy", "scipy"))

# Use any python functions coded in R
my_py_function()
```

